---
title: "R Coding Exercise: Processing, Plotting, and Fitting Models"
output:
  html_document:
    toc: FALSE
---
*Note: To make for a cleaner look, I used {r, message/results/echo/warning = FALSE} to hide the output.*

# Load Libraries

```{r, message=FALSE}
library(dslabs)
library(tidyverse)
library(renv)
```
# Working with Gapminder Data

## Examine Structure

```{r, results= FALSE}
help(gapminder) #Pulls up help page for gapminder data
str(gapminder) #Overview of data structure
summary(gapminder) #Summary of data
class(gapminder) #Determine type of object gap minder is (data frame)
as_tibble(gapminder) #Displays dataframe as a readable tibble 
```

## Condense dataset to *African Countries*

```{r, results = FALSE}
africadata <- gapminder %>%
  filter(continent %in% "Africa")

str(africadata) #Overview of africadata structure
summary(africadata) #Summary of #africadata. This is good for seeing NAs in data
```

## Create two new objects: Infant Mortality (im) and Population (pop)

```{r, results = FALSE}
im <- 
  africadata %>%
  select(infant_mortality, life_expectancy)

pop <- 
  africadata %>%
  select(population, life_expectancy)

str(im)
summary(im)

str(pop)
summary(pop)
```
## Plotting

### Life Expectancy as a function of Infant Mortality

```{r, echo = FALSE, warning = FALSE}
im %>%
  ggplot() +
  geom_point(
    aes(
      x= infant_mortality,
      y= life_expectancy),
    color = "darkgreen") + #Setting data point color
  theme_bw()+ #Adds a nice theme to the figure
  labs(
    x= "Infant Mortality (Infant death per 1000)", #x axis label
    y= "Life Expectancy (Yrs)", #y axis label
    title= "Life Expectancy as a Function of Infant Mortality") + #Title Label
  theme(plot.title = element_text(hjust = 0.5)) #Centers title to middle of figure
```

### Life Expectancy as a function of Population

```{r, echo= FALSE, warning = FALSE}
pop %>%
  ggplot() +
  geom_point(
    aes(
      x= log10(population), #Log scale for Population 
      y= life_expectancy),
    color = "darkred") +
  theme_bw()+
  labs(
    x= "Population (logscale)",
    y= "Life Expectancy (Yrs)",
    title= "Life Expectancy as a Function of Population") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Condense dataset to the year 2000

## Code for seeing which years have NAs for Infant Mortality

```{r, results = FALSE}
im_na<- africadata %>%
  filter(is.na(infant_mortality))
im_na
```

**1960-1981 and 2016 have NAs**

## Create new object that looks at the year 2000

```{r, results = FALSE}
two_k<- africadata %>%
  filter(year %in% "2000")
str(two_k)
summary(two_k)
```

## Make two new objects for the year 2000

```{r}
im2<- two_k %>%
select(infant_mortality, life_expectancy)

pop2<- two_k %>%
  select(population, life_expectancy)
```

## Plotting 

### Life Expectancy as a funtion of Infant Mortality in 2000

```{r, echo = FALSE}
im2 %>%
  ggplot()+
  geom_point(
    aes(
      x = infant_mortality, 
      y = life_expectancy),
    color  = "darkgreen") +
   theme_bw()+
  labs(
    x= "Infant Mortality",
    y= "Life Expectancy (Yrs)",
    title= "Life Expectancy as a Function of Infant Mortality in the Year 2000") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Life Expectancy as a funtion of Population in 2000

```{r, echo = FALSE}
pop2 %>%
  ggplot()+
  geom_point(
    aes(
      x = log10(population), 
      y = life_expectancy),
    color  = "darkred") +
   theme_bw()+
  labs(
    x= "Population (logscale)",
    y= "Life Expectancy(Yrs)",
    title= "Life Expectancy as a Function of Population in the Year 2000") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Linear Models

```{r}
im2.fit <- lm(infant_mortality~life_expectancy, data=im2)

pop2.fit<- lm(population~life_expectancy, data= pop2)

summary(im2.fit)
summary(pop2.fit)
```

**Based on p-values, it looks like there is a significant relationship between infant mortality and life expectancy in the year 2000(p-value = 2.83e-8). There does not seem to be a significant relationship between population and Life expectancy (p-value = 0.62)**
