---
title: "R Coding Exercise: Processing, Plotting, and Fitting Models"
output:
  html_document:
    toc: FALSE
---

*Note: To make for a cleaner look, I used {r, message/results/echo/warning = FALSE} to hide the output.*

# Load Libraries

```{r, message=FALSE}
library(dslabs)
library(tidyverse)
library(renv)
```

# Working with Gapminder Data

## Examine Structure

```{r, results= FALSE}
help(gapminder) #Pulls up help page for gapminder data
str(gapminder) #Overview of data structure
summary(gapminder) #Summary of data
class(gapminder) #Determine type of object gap minder is (data frame)
as_tibble(gapminder) #Displays dataframe as a readable tibble 
```

## Condense dataset to *African Countries*

```{r, results = FALSE}
africadata <- gapminder %>%
  filter(continent %in% "Africa")

str(africadata) #Overview of africadata structure
summary(africadata) #Summary of #africadata. This is good for seeing NAs in data
```

## Create two new objects: Infant Mortality (im) and Population (pop)

```{r, results = FALSE}
im <- 
  africadata %>%
  select(infant_mortality, life_expectancy)

pop <- 
  africadata %>%
  select(population, life_expectancy)

str(im)
summary(im)

str(pop)
summary(pop)
```

## Plotting

### Life Expectancy as a function of Infant Mortality

```{r, echo = FALSE, warning = FALSE}
im %>%
  ggplot() +
  geom_point(
    aes(
      x= infant_mortality,
      y= life_expectancy),
    color = "darkgreen") + #Setting data point color
  theme_bw()+ #Adds a nice theme to the figure
  labs(
    x= "Infant Mortality (Infant death per 1000)", #x axis label
    y= "Life Expectancy (Yrs)", #y axis label
    title= "Life Expectancy as a Function of Infant Mortality") + #Title Label
  theme(plot.title = element_text(hjust = 0.5)) #Centers title to middle of figure
```

### Life Expectancy as a function of Population

```{r, echo= FALSE, warning = FALSE}
pop %>%
  ggplot() +
  geom_point(
    aes(
      x= log10(population), #Log scale for Population 
      y= life_expectancy),
    color = "darkred") +
  theme_bw()+
  labs(
    x= "Population (logscale)",
    y= "Life Expectancy (Yrs)",
    title= "Life Expectancy as a Function of Population") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Condense dataset to the year 2000

## Code for seeing which years have NAs for Infant Mortality

```{r, results = FALSE}
im_na<- africadata %>%
  filter(is.na(infant_mortality))
im_na
```

**1960-1981 and 2016 have NAs**

## Create new object that looks at the year 2000

```{r, results = FALSE}
two_k<- africadata %>%
  filter(year %in% "2000")
str(two_k)
summary(two_k)
```

## Make two new objects for the year 2000

```{r}
im2<- two_k %>%
select(infant_mortality, life_expectancy)

pop2<- two_k %>%
  select(population, life_expectancy)
```

## Plotting

### Life Expectancy as a funtion of Infant Mortality in 2000

```{r, echo = FALSE}
im2 %>%
  ggplot()+
  geom_point(
    aes(
      x = infant_mortality, 
      y = life_expectancy),
    color  = "darkgreen") +
   theme_bw()+
  labs(
    x= "Infant Mortality",
    y= "Life Expectancy (Yrs)",
    title= "Life Expectancy as a Function of Infant Mortality in the Year 2000") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Life Expectancy as a funtion of Population in 2000

```{r, echo = FALSE}
pop2 %>%
  ggplot()+
  geom_point(
    aes(
      x = log10(population), 
      y = life_expectancy),
    color  = "darkred") +
   theme_bw()+
  labs(
    x= "Population (logscale)",
    y= "Life Expectancy(Yrs)",
    title= "Life Expectancy as a Function of Population in the Year 2000") +
  theme(plot.title = element_text(hjust = 0.5))
```

# Linear Models

```{r}
im2.fit <- lm(infant_mortality~life_expectancy, data=im2)

pop2.fit<- lm(population~life_expectancy, data= pop2)

summary(im2.fit)
summary(pop2.fit)
```

**Based on p-values, it looks like there is a significant relationship between infant mortality and life expectancy in the year 2000(p-value = 2.83e-8). There does not seem to be a significant correlation between population and life expectancy (p-value = 0.62)**

# *This section was added by Kimberly Perez*

```{r}
library(broom)
```


# **Region and Fertility**

```{r}
# Code to create a box plot to visualize fertility by region
bp_fert<-ggplot(two_k, aes(x=region, y=fertility, color= region)) + geom_boxplot(outlier.color="seagreen3", outlier.shape=8, outlier.size=2)

# Editing the size and shapes present on the boxplot and adding title/axis labels 
bp_fert + stat_summary(fun=mean, geom="point", shape=13, size=4) + labs(x="Region", y= "Fertility (Number of children)", color="Region Legend", title="Fertility by Region in Africa for 2000")
```

# **Filtering Dataset for NAs (fertility and GDP) and condensing data to year 2000**

```{r, results= FALSE}
fert_na<- africadata %>%
  filter(is.na(fertility))
fert_na
unique(fert_na$year)

gdp_na<- africadata %>%
  filter(is.na(gdp))
gdp_na
unique(gdp_na$year)

fert_a<- africadata[which(africadata$year=="2000"),]

str(fert_a)
```

# **Fertility as a function of Gross Domestic Production (GDP) in 2000**

```{r}
fert_a %>%
  ggplot() +
  geom_point(
    aes(
      x= log10(gdp), #Log scale for Population 
      y= fertility),
    color = "cadetblue") +
  theme_bw()+
  labs(
    x= "GDP (logscale)",
    y= "Fertility (Number of Children)",
    title= "Fertility as a Function of Gross Domestic Production (GDP) in 2000") +
  theme(plot.title = element_text(hjust = 0.5))
```

# **Fertility as a function of Gross Domestic Production (GDP)**

```{r, warning=FALSE}
ggplot(fert_a, aes(log10(gdp), fertility))+ geom_point() +
  geom_smooth(method="lm") +
  labs(title="Fertility as a function of Gross Domestic Production", x="Gross Domestic Product [GDP] (logscale)", y="Fertility (Number of children)")
```

# **A Simple Fit and Table Using the Package Broom**
```{r}
le_fit<-lm(gdp~life_expectancy, data=two_k)
tidy(le_fit)
```

Based on the p-values for the given fits, GDP as a predictor of LE is said to be statistically significant. But can we truly trust p-values?…

