---
title: "Tidy Tuesday Exercise2"
output: 
  html_document:
    toc: FALSE
---

# Getting Started

#### Libraries

```{r}
library(tidyverse)
library(here)
library(rsample) #Data spliting
library(tidymodels) #Modeling 
```

#### Load Data

```{r}
cage<- read_csv(here("data", "cage-free-percentages.csv")) #Cage free %
egg<- read_csv(here("data", "egg-production.csv"))
```

#### Look at Data sets

```{r}
head(cage)
head(egg)
```

Note: Within the `egg` data set, the value 'all' includes cage-free and conventional housing.

# Initial Cleaning

#### Cage Data set

```{r}
cage<- 
  cage %>%
  select(!source) #Removes Source column
```

#### Egg Data set

```{r}
egg<- 
  egg %>%
  select(!source) %>% #Removes Source Column
  mutate(egg_hen = n_eggs / n_hens, #Creates column that gives an average eggs per hen
         prod_process = recode(prod_process, "cage-free (non-organic)" = "cage_free_no"),
         prod_process = recode(prod_process, "cage-free (organic)" = "cage_free_o"))


```

# Exploratory Visualization

#### Fig. 1: Percentage of cage free gens and eggs

```{r}
cage %>%
  ggplot() + geom_line( #Percent of cage free hens (orange)
    aes(x = observed_month,
        y = percent_hens),
    color = "#f68f3c") +
  geom_point(
    aes(x = observed_month,
        y = percent_hens),
    size = 1,
    color = "#f68f3c")+
  geom_line( #Percent of cage free eggs (green)
    aes(x = observed_month,
        y = percent_eggs),
    color = "#5e8d5a",
    size = 1) +
  theme_bw() +
  labs(x = "Year",
       y = "%",
       title = "Fig.1 :Percentage of Cage Free Hens and Eggs over Time") 
  
```

Percentage of Hens are in orange, and percentage of eggs are in green. We see an increase in percentage of cage free eggs and hens over time.

#### Fig. 2: Eggs per Hen across production type and process

```{r}
egg %>%
  ggplot() + geom_boxplot(
    aes(x = prod_type,
        y = egg_hen,
        color = prod_process)) +
  theme_bw()+
  labs(title = "Fig. 2: Eggs per Hen Across Production Process and Type",
       x = "Production Type",
       y = "Eggs per Hen",
       color = "Production Process") + 
  scale_color_manual(values = c(all = "#ee6f68",
                                cage_free_no = "#5e8d5a",
                                cage_free_o = "#f68f3c")) 
```

Figure 2 depicts a stark difference in the number eggs per hen between hatchilng eggs and table eggs. No difference is observed between the production processes within the table eggs.

#### Fig. 3: Eggs per hen over time

```{r}
egg %>%
  ggplot() + geom_line(
    aes(x = observed_month,
        y = egg_hen,
        color = prod_type)) +
  theme_bw()+
  labs(x = "Year",
       y = "Eggs per Hen",
       title = "Fig. 3: Eggs Per hen over Time by Production Type",
       color = "Production Type")
```

Similarly, Figure 3 shows that more eggs per hen are produced from the table eggs process.

### Question/Hypothesis:

Because a higher portion of the population consumes eggs as opposed to hatching chicks, and due to the energetically taxing/limiting process of hatching, we expect to see a significantly higher effort towards the production of table eggs as opposed to hatching eggs.

-   Predictor: Production Type (Hatching or Table Eggs)

-   Outcome: Eggs per hen

# Final Cleaning

We will use only the egg dataset.

#### Remove Columns that are not needed for analysis

```{r}
egg <- 
  egg %>%
  select(prod_type, egg_hen)
```

#### Split into Test and Train Set

```{r}
# Fix the random numbers by setting the seed 
# This enables the analysis to be reproducible when random numbers are used 
set.seed(222)
# Put 3/4 of the data into the training set 
data_split <- initial_split(egg, prop = 3/4)

# Create data frames for the two sets:
train <- training(data_split)
test  <- testing(data_split)
```

# Modeling

## Null Model

#### 5-Fold Cross Validation

```{r}
fold_egg_train <- vfold_cv(train, v = 5, repeats = 5, strata = egg_hen)

fold_egg_test <- vfold_cv(test, v = 5, repeats = 5, strata = egg_hen)
```

#### Recipes

```{r}
#Train Data
egg_rec_train <- 
  recipe(egg_hen ~ prod_type, data = train) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_ordinalscore() %>%
  step_zv(all_predictors()) 


#Test Data
egg_rec_test <- 
  recipe(egg_hen ~ prod_type, data = train) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_ordinalscore() %>%
  step_zv(all_predictors()) 

```

#### Define Model

```{r}
lm_mod <- 
  linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")

```

#### Train Data

##### Recipe

```{r}
null_train_rec<- 
recipe(egg_hen ~ 1, data = train)

null_train_rec
```

##### Workflow

```{r}
null_train_wf <- 
  workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(null_train_rec)
```

##### Fit

```{r, results=FALSE}
null_train_fit <- 
  fit_resamples(null_train_wf, resamples = fold_egg_train)
```

##### RMSE Metric

```{r}
null_train_met <- collect_metrics(null_train_fit)

null_train_met
```

**RMSE = 2.18 and STD_ERR = 0.011**

#### Test Data

##### Recipe

```{r}
null_test_rec<- 
  recipe(egg_hen ~ 1, data = test)
```

##### Workflow

```{r}
null_test_wf <- 
  workflow() %>% 
  add_model(lm_mod) %>% 
  add_recipe(null_test_rec)
```

##### Fit

```{r}
null_test_fit <- 
  fit_resamples(null_test_wf, resamples = fold_egg_test)
```

##### RMSE Metric

```{r}
null_test_met <- collect_metrics(null_test_fit)

null_test_met
```

**RMSE = 2.19 and STD_ERR = 0.084\
**

## Tree Model

#### Model Specification

```{r}
tune_spec_dtree <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()) %>%
  set_engine("rpart") %>% 
  set_mode("regression")

tune_spec_dtree
```

Think of `tune()` here as a placeholder. After the tuning process, we will select a single numeric value for each of these hyperparameters. For now, we specify our parsnip model object and identify the hyperparameters we will `tune()`.

#### Workflow Definition

```{r}
dtree_wf <- workflow() %>%
  add_model(tune_spec_dtree) %>%
  add_recipe(egg_rec_train)
```

#### Tuning Grid Specification

We can create a regular grid of values to try using some convenience functions for each hyperparameter:

```{r}
#create a regular grid of values for using convenience functions for each hyperparameter.
tree_grid_dtree <-
  grid_regular(
    cost_complexity(), 
    tree_depth(), 
    levels = 5)

tree_grid_dtree
```

#### Tuning using Cross-validation and `tune_grid()` function

```{r}
dtree_resample <- 
  dtree_wf %>% 
  tune_grid(
    resamples = fold_egg_train,
    grid = tree_grid_dtree)
    
```

Once we have our tuning results, we can both explore them through visualization and then select the best result. The function `collect_metrics()` gives us a tidy tibble with all the results

```{r}
dtree_resample %>%
  collect_metrics()
```

##### Plot Model using autoplot()

```{r}
dtree_resample %>%
  collect_metrics() %>%
  mutate(tree_depth = factor(tree_depth)) %>%
  ggplot(aes(cost_complexity, mean, color = tree_depth)) +
  geom_line(linewidth = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_color_viridis_d(option = "plasma", begin = .9, end = 0)
```

<!--# SHould it look like this? -->

##### Show and Select best performing models

The [`show_best()`](https://tune.tidymodels.org/reference/show_best.html) function shows us the top 5 candidate models by default. We set n=1

```{r}
dtree_resample %>%
  show_best(n=1)
```

<!--# Update -->**From the plot and the tibble above, we see that the the model with treedepth =1 has the lowest rmse value with a mean of 1.19 and standard error of 0.018.**

We can also use the [`select_best()`](https://tune.tidymodels.org/reference/show_best.html) function to pull out the single set of hyperparameter values for our best decision tree model:

```{r}
#Selects best performing model
best_tree <- dtree_resample %>%
  select_best()

best_tree
```

##### Create final fit based on best model permutation and plotting predicted values from that final fit model

We can update (or "finalize") our workflow object `tree_wf` with the values from `select_best()`.

```{r}
dtree_final_wf <- 
  dtree_wf %>% 
  finalize_workflow(best_tree)

dtree_final_wf
```

```{r}
#Create workflow for fitting model to train predictions
dtree_final_fit <- 
  dtree_final_wf %>%
  fit(train) 
```

##### Calculating Residuals and Plotting Actual Vs. Predicted Values

```{r}
dtree_residuals <- dtree_final_fit %>%
  augment(train) %>% #use augment() to make predictions from train data
  select(c(.pred, egg_hen)) %>%
  mutate(.resid = egg_hen - .pred) #calculate residuals and make new row.

dtree_residuals
```

##### Predictions vs. Actual

```{r}
dtree_pred_plot <- ggplot(dtree_residuals, 
                          aes(x = egg_hen, 
                              y = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Actual: Decision Tree", 
       x = "Egg_hen Outcome", 
       y = "Egg_hen Prediction")
dtree_pred_plot
```

##### Predictions vs. Residuals

```{r}
dtree_residual_plot <- ggplot(dtree_residuals, 
                              aes(y = .resid, 
                                  x = .pred)) + 
  geom_point() + 
  labs(title = "Predictions vs Residuals: Decision Tree", 
       x = "Egg_hen Prediction", 
       y = "Residuals")
plot(dtree_residual_plot) #view plot
```

## 

## Random Forest Model

## Lasso Model

<!--# Need a 4th  -->

Fit at least 4 different ML models to the data using the `tidymodels` framework we practiced. Use the CV approach for model training/fitting. Explore the quality of each model by looking at performance, residuals, uncertainty, etc. All of this should still be evaluated using the training/CV data. You can of course recycle code from the previous exercise, but I also encourage you to explore further, e.g. try different ML models or use different metrics. You might have to do that anyway, depending on your question/outcome.

Based on the model evaluations, decide on one model you think is overall best. Explain why. It doesn't have to be the model with the best performance. You make the choice, just explain why you picked the one you picked.

#### Final Assessment

Based on the model evaluations, decide on one model you think is overall best. Explain why. It doesn't have to be the model with the best performance. You make the choice, just explain why you picked the one you picked.

# Discussion
