---
title: "exploration"
format:
  html:
    theme: default
---

## Library

```{r}
library(tidyverse) #For plotting and wrangling
library(here) #Setting working directory 
library(gtsummary) #For summary tables 
```

## Load Data

```{r}
d<- readRDS(here("fluanalysis", "data", "SypAct_clean.rds"))
```

## Produce and Print a Summary Table for both Body Temperature and Nausea

#### Summary Stats

```{r}
summary(d$BodyTemp)
summary(d$Nausea)
```

#### Summary of full data set

```{r}
tab<- 
  tbl_summary(d)
tab
```

## Create Histogram for Body Temperature

```{r}
d %>%
  ggplot()+geom_histogram(
    aes(
      x = BodyTemp))
```

**A majority of the body temperatures fall between 98-99°F**

## Let's look at Predictor Variables for our outcomes of interest (Body Temperature and Nausea)

### Body Temperature

#### Weakness and Body Temp

```{r}
d %>%
  ggplot()+geom_boxplot(
    aes(
      x= Weakness,
      y = BodyTemp))
```

#### Chills and Sweats and Body Temp

```{r}
d %>%
  ggplot()+geom_boxplot(
    aes(
      x= ChillsSweats,
      y = BodyTemp))
```

#### Fatigue and Body Temp

```{r}
d %>%
  ggplot()+geom_boxplot(
    aes(
      x= Fatigue,
      y = BodyTemp))
```

#### Fever and Body Temp

```{r}
d %>%
  ggplot()+geom_boxplot(
    aes(
      x= SubjectiveFever,
      y = BodyTemp))
```

**On average, it looks like those that experienced severe weakness symptoms, chills and sweats, fatigue, and fever had higher body temperatures. We would expect to see this positive relationship in the latter- so fever will be our main predictor of interest for Body Temperature.**

## Now let's look at some predictor variables for the categorical variable (Nausea)

#### Diarrhea and Nausea

```{r}
d %>%
  ggplot() + geom_count(
    aes(
      x = Nausea,
      y = Fatigue))
```

#### Chills and Nausea

```{r}
d %>%
  ggplot() + geom_count(
    aes(
      x = Nausea,
      y = ChillsSweats))
```

#### Fatigue and Nausea

```{r}
d %>%
  ggplot() + geom_count(
    aes(
      x = Nausea,
      y = Fatigue))
```

#### Fever and Nausea

```{r}
d %>%
  ggplot() + geom_count(
    aes(
      x = Nausea,
      y = SubjectiveFever))
```

**Fever, Fatigue, Chills, and Diarrhea all appear to have equal or no major positive relationship (Where x:nausea = Yes-Yes \< Yes-No). Vomiting will be our predictor of interest for the categorical outcome (Nausea), as there was a large proportion that did not experience either Nausea or Vomiting, and a medium-proportion of those that experienced one symptom with the other.**
